---
title: "2020-07 Employment Inventory 2018"
author: "Purva Singh & Kelsie Telson"
date: "7/24/2020"
output: html_document
---

The following report contains high-level results from QC testing conducted on the 2018 Employment Inventory dataset. The review was conducted by QA analysts Purva Singh and Kelsie Telson in July 2020. Due to the confidentiality of EDD data, only aggregate information will be displayed here. To see the full Test Plan associated with this report, including select record-level findings, please see "R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2019\QAQC\Office of QA\Test Plan 2018 Employment Inventory.xlsx".

```{r setup, include=FALSE}
pkgTest <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dep = TRUE)
  sapply(pkg, require, character.only = TRUE)}


connect_datawarehouse <- function() {
  
  # Connect to Datahub using RODBC package
  channel <- RODBC::odbcDriverConnect(
    paste0("driver={SQL Server}; server=sql2014b8;
             database=EMPCORE;
             trusted_connection=true"))
  
  # Return open RODBC connection
  return(channel)
}



packages <- c("RODBC","tidyverse","openxlsx","hash","readtext","data.table","dplyr", "knitr")
pkgTest(packages)


# connect to database
channel <- connect_datawarehouse()

#retrieve data from sql server
#note: variables [SHAPE] and [GDB_GEOMATTR_DATA] not included due to memory restrictions
raw_dt <- data.table::as.data.table(
  RODBC::sqlQuery(channel,
                  paste0("
SELECT [OBJECTID]
      ,[Status]
      ,[Score]
      ,[Match_type]
      ,[Match_addr]
      ,[emp_id]
      ,[dba]
      ,[address]
      ,[city]
      ,[zip]
      ,[emp1]
      ,[emp2]
      ,[emp3]
      ,[payroll]
      ,[naics]
      ,[own]
      ,[meei]
      ,[init]
      ,[end_]
      ,[react]
      ,[Run]
      ,[Check_]
      ,[flag]
      ,[Move]
      ,[Comment]
      ,[MGRA13]
  FROM [EMPCORE].[dbo].[CA_EDD_EMP2018]"),
                  stringsAsFactors = FALSE),
  stringsAsFactors = FALSE)


#note: variables [SHAPE] and [GDB_GEOMATTR_DATA] not included due to memory restrictions
raw_dt_hqd <- data.table::as.data.table(
  RODBC::sqlQuery(channel,
                  paste0("SELECT [OBJECTID]
      ,[emp_id]
      ,[sub_emp_id]
      ,[dba]
      ,[Comment]
      ,[Check_]
      ,[share]
      ,[MGRA13]
  FROM [EMPCORE].[dbo].[CA_EDD_EMP2018HQD]"),
                  stringsAsFactors = FALSE),
  stringsAsFactors = FALSE)

```

## Test 1

1c. Check for no NaNs, 0s in dataset
The majority of variables have NAs  across multiple records, however there are no NA values for emp_id or MGRA13. Blake notes the NAs are acceptable as long as they are not in these variables. 

There are 2 records for which [ZIP]= 0. These records should be reviewed by the requesting team.
Additionally, variables [emp1], [emp2], [emp3], and [payroll] display a number of records equal to 0. These cases were investigated in more detail in Test 2.

1d. Check datatypes of each column makes sense
All variable types appear to be appropriate.

```{r test1}
summary(raw_dt)

test1c_na<-as.data.table(raw_dt %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.)))))

test1c_zeros<-rbind(table(raw_dt$zip==0),
      table(raw_dt$emp1==0),
      table(raw_dt$emp2==0),
      table(raw_dt$emp3==0),
      table(raw_dt$payroll==0))
test1c_zeros<- as.data.table(cbind(variable=c("zip", "emp1", "emp2", "emp3", "payroll"),
                                   test1c_zeros))
setnames(test1c_zeros, old = c('FALSE','TRUE'), new = c('Not Zero','Zero'))

kable(test1c_na, caption="Test 1c: Records with NA")

kable(test1c_zeros, caption="Test 1d: Records with 0")

```

## Test 4

4c. 'own' should be discretely distributed, no trend
Nearly all records have assignment of "5"

4d. 'meei' should be discretely distributed, no trend
Most records assigned "1", followed by "3", approximately even distribution of "4", "5", and "6". There are no records with "2".

The PowerBI file containing the visuals associated with this analysis may be found at
"R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2019\OutputData\2018 Employment Inventory QA.pbix"


## Test 7

7a. Check totals against BLS Quarter 3 2018 (establishments, employees)
Totals in the SANDAG dataset are slightly lower (~23,000-24,000) for emp1, emp2, and emp3 as compared to BLS. Payroll is approximately $449 million less than BLS. 

7b. Check totals against published annual totals from EDD Organization
"R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2018\SourceData"
Analysis in PowerBI and R revealed discrepancies between txt files and final data from the SQL server for [emp1], [emp2], [emp3], and [payroll] variables. Overall, the final data files show less records than the raw txt files. 
**This finding requires additional investigation from the requesting team.**



```{r test7}
#Test 7
test7a<- as.data.table(cbind(variable=c("emp1","emp2", "emp3", "payroll"),
              rbind(sum(raw_dt$emp1, na.rm=TRUE),
                    sum(raw_dt$emp2, na.rm=TRUE),
                    sum(raw_dt$emp3, na.rm=TRUE),
                    sum(raw_dt$payroll, na.rm=TRUE)),
              bls=c(1458131,1461622,1466234,21848394370))) #values found on bls website, screenshot in project folder

test7a$V2<- as.numeric(test7$V2)
test7a$bls<- as.numeric(test7$bls)
test7a$diff<- test7$V2-test7$bls

kable(test7a, caption="Test 7a: BLS Comparison")

```

## Test 9

9a. Employment by month per company should be consistent and shouldn't show any major changes, barring a change in operation state of the company

A visual review was done for differences between Month 1 and Month 2, Month 2 and Month 3, and Month 1 and Month 3 (net difference over quarter). Overall, employee change over time was relatively stable for most employers. 

There were several notable exceptions where differences appeared dramatic during a visual review.

The PowerBI file containing the visuals associated with this analysis may be found at
"R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2019\OutputData\2018 Employment Inventory QA.pbix"

##Other Findings

a. Input in variables [end_] and [react] for several records appears invalid. See Test Plan for details. The PowerBI file containing the visuals associated with this analysis may be found at
"R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2019\OutputData\2018 Employment Inventory QA.pbix"

b. There was a strong positive relationship between the number of [emp1]/[emp2]/[emp3, and the [payroll] for most employers. One employer was an outlier, having a higher payroll and lower employee count than most other employers. Additionally, one employer had a much higher payroll and employee count than most of the other employers in the region, however the payroll-employee relationship for this employer is still aligned with the overall trend. The PowerBI file containing the visuals associated with this analysis may be found at
"R:\DPOE\Employment\CAEDD\QCEW Microdata\RESTRICTED\2019\OutputData\2018 Employment Inventory QA.pbix"

